---
title: "Exploration of NFHS death file"
author: Casey Breen
---


## Read in data 


```{r}
## library packages 
library(tidyverse)
library(haven)
library(cowplot)
library(here)
library(srvyr)

## read in death file 
death_file <- read_dta(here("data", "analysis_files", "nfhs5_deaths.DTA"))
```


## clean up variables 

- Each row corresponds to a death 
- Each column corresponds to a variable


variable definitions to recode: 

- hv024 = state of household (new var, state)
- shdist = district of household (new var, district)
- sh91_ = death_registration (new var, registered)
- hv001 = cluster number 
- hv005 = household weight 
- hv009 = number of household members

other key variables (not recodes)

- death age = age of death 
- female = female gender 
- native_language = native language 
- altitude_quintile = altitude quintile 
- pregnancy_death = pregnancy related death 
- external_death = external cause of death
- month_since_death = months since death
- caste_religion = caste religion 
- highest_education = highest education of household member 
- working_members = working age household members 

```{r}
## death registration 
## 1 = registered and 0 = not registered
death_file <- death_file %>% 
  mutate(registered_string = case_when(
    registered == 1 ~ "registered",
    registered == 0 ~ "not registered"
  )) %>% 
  ## recode gender 
  mutate(gender = case_when(
    female == 1 ~ "women",
    TRUE ~ "men"
  )) 

```


```{r}
## read in sryvr package 
death_survey <- death_file %>%
  as_survey_design(weights = weight, ids = cluster)
```


## geography 

```{r}
## death registration by district 
death_registration_by_district <- death_survey %>% 
  group_by(district, gender) %>% 
  summarize(mean_death_registration = survey_mean(registered, vartype = "se"))

# Create the distribution plot
district_plot <- death_registration_by_district %>% 
  ggplot(aes(x = mean_death_registration, color = gender, fill = gender)) + 
  geom_density(aes(y = after_stat(count) / sum(after_stat(count))), alpha = 0.2) +  # Adjust allows for smoothing
  cowplot::theme_cowplot() + 
  labs(x = "Deaths Registered (%)", y = "Proportion", color = "Gender", fill = "Gender") + 
  scale_x_continuous(labels = scales::percent_format()) +
  ggsci::scale_fill_lancet() +
  ggsci::scale_color_lancet() + 
  theme(legend.position = "bottom")
```


```{r}
# Define burnt orange color
burnt_orange <- "#CC5500"

# Calculating mean death registration and standard error by state
death_registration_by_state_new <- death_survey %>% 
  group_by(state, gender) %>% 
  summarize(mean_death_registration = survey_mean(registered, vartype = "se")) %>% 
      mutate(state = as_factor(state)) %>% 
  mutate(state = case_when(
    state == "dadra & nagar haveli and daman & diu" ~ "DNH & DD",
    TRUE ~ state))

# Plotting with geom_linerange for 80% and 90% CI and geom_pointrange for 95% CI
state_plot <- death_registration_by_state_new %>% 
  ggplot() +
  geom_linerange(aes(x = mean_death_registration, 
                     y = reorder(state, mean_death_registration),
                     xmin = mean_death_registration - 1.645 * mean_death_registration_se,
                     xmax = mean_death_registration + 1.645 * mean_death_registration_se,
                     color = gender), 
                 size = 1,
                 position = position_dodge(0.4)) +
  geom_pointrange(aes(x = mean_death_registration, 
                      y = reorder(state, mean_death_registration),
                      xmin = mean_death_registration - 1.96 * mean_death_registration_se,
                      xmax = mean_death_registration + 1.96 * mean_death_registration_se, 
                      color = gender), 
                  fill = burnt_orange,
                  size = 1.5, 
                  fatten = 2.5,
                  position= position_dodge(0.4)) + 
  cowplot::theme_cowplot() + 
  ggsci::scale_color_lancet() +
  labs(x = "Deaths Registered (%)",
       y = "State") + 
  scale_x_continuous(labels = scales::percent_format(), limits = c(0, 1)) + 
  theme(legend.position = "bottom") 
```


```{r}
## age plot df 
age_plot_df <- death_survey %>% 
  mutate(
    female_recode = case_when(
      female == 1 ~ "women",
      TRUE ~ "men"
    ),
    # Create 5-year age groups
    age_group = cut(death_age, breaks = seq(0, max(death_age, na.rm = TRUE), by = 5), right = FALSE, include.lowest = TRUE)
  ) %>%
  group_by(age_group, female_recode) %>% 
  summarize(registered_mean = survey_mean(registered, vartype = "se")) 
  
## age plot 
age_plot <- ggplot(age_plot_df, aes(x = age_group, y = registered_mean, color = female_recode,
             ymin = registered_mean - 1.96 * registered_mean_se, 
             ymax = registered_mean + 1.96 * registered_mean_se)) +
  geom_pointrange(position = position_dodge(width = 0.4)) + 
  geom_line(position = position_dodge(width = 0.4)) + 
  ggsci::scale_color_lancet() +
  theme_cowplot() + 
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 45, hjust = 1)) + # Rotating the x-axis labels
  labs(x = "Age of Death",
       y = "Death Registration") + 
  guides(color = guide_legend(title = NULL))

```

```{r}
urban_plot <- death_survey %>% 
  mutate(rural = case_when(
    rural == 1 ~ "rural",
    TRUE ~ "urban"
  )) %>%
  group_by(wealth_quintile, gender, rural) %>% 
  summarize(registered = survey_mean(registered, vartype = "se")) %>% 
  ggplot(aes(x = wealth_quintile, y = registered)) + 
  geom_col(aes(fill = gender), color = "black", position = position_dodge2(width = 1, padding = 0.2), alpha = 0.5) + 
  geom_text(aes(label = paste0(round(registered, 2)* 100, "%")), vjust = -0.8, hjust = .4, position = position_dodge2(width = 1, padding = 0.2)) + 
  ggsci::scale_fill_lancet() +
  theme_cowplot() + 
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  theme(legend.position = "bottom") + 
  labs(x = "Wealth Quantile",
       y = "Death Registration") + 
  facet_wrap(~rural)
  
```

```{r}
plot_left <- cowplot::plot_grid(district_plot, urban_plot, age_plot, labels = "AUTO", cols = 1) 

full_plot <- cowplot::plot_grid(plot_left, state_plot, labels = c("", "D"), cols = 2, rel_widths = c(1.3, 1))

ggsave(full_plot, filename = here("figures", "death_registration.png"), width = 14, height = 10, bg = "white")
```




```{r}
death_file %>% 
  filter(facility_distance < 100) %>% 
  group_by(registered) %>% 
  summarize(mean(facility_distance))


death_file %>% 
  group_by(registered) %>% 
  summarize(mean(facility_distance))

death_file %>% 
  mutate(distance = ntile(road_length_km/area_km2, n = 5)) %>% 
  group_by(distance) %>% 
  summarize(mean(registered))


death_file %>% 
  mutate(facility_distance_quantile = ntile(facility_distance, n = 5)) %>% 
  group_by(facility_distance_quantile) %>% 
  summarize(facility_distance = mean(registered)) %>% 
  ggplot()




death_file %>% 
 # count(is.na(road_length_km)) 
  mutate(distance = ntile(road_length_km, n = 5)) %>% 
  group_by(distance) %>% 
  summarize(mean = mean(registered))


```



```{r}
# First figure: Facility distance quantile vs mean registration likelihood
facility_distance_plot <- death_file %>% 
  mutate(facility_distance_quantile = ntile(facility_distance, n = 5)) %>% 
  group_by(facility_distance_quantile) %>% 
  summarize(mean_registered = mean(registered, na.rm = TRUE)) %>% 
  ggplot(aes(x = facility_distance_quantile, y = mean_registered)) +
  geom_line() + 
  geom_point() + 
  labs(x = "Facility Distance (Quantile)", y = "Death Registration (%)", 
       title = "Facility Distance") +
  theme_cowplot()

# Second figure: Road length quantile vs mean registration likelihood
road_length_plot <- death_file %>% 
  mutate(road_length_quantile = ntile(road_length_km/area_km2, n = 5)) %>% 
  group_by(road_length_quantile) %>% 
  summarize(mean_registered = mean(registered, na.rm = TRUE)) %>% 
  ggplot(aes(x = road_length_quantile, y = mean_registered)) +
  geom_line() + 
  geom_point() + 
  labs(x = "Road Length/km^2 (Quantile)", y = "Death Registration (%)", 
       title = "Road Length") +
  theme_cowplot()

# Display both plots
death_registration_distance <- plot_grid(facility_distance_plot, road_length_plot)

ggsave(death_registration_distance, filename = here("figures", "death_registration_distance.png"), width = 8, height = 4, bg = "white")
```




```{r}
death_reg_over_time <- death_file %>% 
  group_by(months_since_death) %>% 
  summarize(registered = mean(registered)) %>% 
  ggplot(aes(x = months_since_death, y = registered)) + 
  geom_line() + 
  geom_point() + 
  theme_cowplot() + 
  labs(x = "Months since death",
       y = "Deaths Registered (%)") + 
  ylim(0, 1)

ggsave(death_reg_over_time, filename = here("figures", "death_reg_over_time.png"), width = 6, height = 4, bg = "white")

```


