---
title: "Machine learning set up"
author: Casey Breen
---

Other predictors: 

— Presence of educated family member
— Household size 
— Geospatial stuff
— Length of period between death and interview (months)
— Caste 
— Native language 
— Number of children under 5
— Marital status 
— Mode of transport 
— Exposure to media and literacy 
— Household member disability 
— District level covariates - district wealth, district mortality, education, etc. etc. 
— Variables for health care access 
— Digital variables - internet access, mobile phone access, etc. 


## notes for aashish 

- Missing in household and missing in 


```{r}
## library packages 
library(tidyverse)
library(haven)
library(cowplot)
library(here)
library(srvyr)
library(sl3)
library(tictoc)
library(yardstick)
library(tidymodels)

## read in death file 
death_file <- read_dta(here("data", "analysis_files", "nfhs5_deaths.DTA")) %>% 
  filter(months_since_death > 3) %>% 
  mutate(registered = as.factor(registered))
```



```{r}
death_file <- death_file %>% 
  mutate(facility_distance_quantile = ntile(facility_distance, n = 5)) %>% 
  mutate(road_length_km_quantile = ntile(road_length_km, n = 5))
  
death_file <- death_file %>% 
  filter(!is.na(area_km2))
```




```{r}
set.seed(12)  # Setting seed for reproducibility

## sample splits 
death_file_split <- initial_split(death_file, 0.75)

training_sample <- training(death_file_split)
test_sample  <- testing(death_file_split)
```




```{r}
## baseline model 
model <- glm(registered ~ death_age + female + as.factor(wealth_quintile) + native_language + as.factor(internet) + highest_education + as.factor(health_insurance) + as.factor(rural) + as.factor(state) + own_land + cycle + external_death + female_head + facility_distance_quantile + road_length_km_quantile + as.factor(altitude_quintile), data = training_sample, family = "binomial")

predict(model, newdata = test_sample, type = "response")

test_sample <- test_sample %>% 
  mutate(prediction_logit = predict(model, newdata = test_sample, type = "response")) 

test_sample %>% 
  mutate(prediction_binary = case_when(
    prediction_logit > 0.5 ~ 1,
    TRUE ~ 0
  )) %>% 
  count(registered, prediction_binary) %>% 
  mutate(prop = n / sum(n))
```




```{r}
# set up 10 fold CV
folds_10fold <- vfold_cv(death_file, 10)

# Define the model specification
rf_model <- rand_forest(trees = 500, min_n = 5, mode = "classification") %>%
  set_engine("ranger", importance = "permutation")

# Define the recipe for data preprocessing
rf_recipe <- recipe(registered ~ death_age + female +
                      wealth_quintile + facility_distance_quantile + region + 
                      road_length_km_quantile + 
                      native_language + internet + 
                      highest_education + health_insurance + 
                      rural + own_land + cycle + 
                      external_death + female_head + state + 
                      altitude_quintile, data = death_file) %>%
  step_dummy(all_nominal_predictors()) %>% # Convert categorical to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize numeric predictors

# Create a workflow
rf_workflow <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(rf_recipe)

# Fit the model
rf_fit <- rf_workflow %>%
  fit_resamples(resamples = folds_10fold,
                control = control_resamples(save_pred = TRUE))

## get error metrics  
error_metrics_loco_cv <- collect_metrics(rf_fit) 

## get error metrics  
predictions <- collect_predictions(rf_fit) 

death_file <- death_file %>% 
  mutate(row = row_number())

predictions <- predictions %>% 
  left_join(death_file, by = c(".row" = "row", "registered")) %>% 
    mutate(registered = as.numeric(registered) - 1)

```


## Predicted observed plot 

```{r}
predicted_observed <- predictions %>% 
  group_by(female) %>% 
  mutate(decile = ntile(.pred_1, 100)) %>% 
  group_by(decile) %>% 
  summarize(reg = mean(registered),
            pred = mean(.pred_1)) %>% 
  ggplot(aes(x = reg, y = pred)) + 
  geom_point() + 
  geom_abline(color = "red", linetype = "dashed") + 
  ylim(0, 1) + 
  xlim(0, 1) + 
  theme_cowplot() + 
  labs(x = "Observed", y = "Predicted")
```


## Make confusion matrix

```{r}
# Create confusion matrix
confusion_matrix <- predictions %>%
  count(registered, .pred_class) %>%
  mutate(prop = paste0(round(n / sum(n) * 100), "%"))


# Plot with corrected code
confusion_matrix_plot <- confusion_matrix %>% 
  mutate(registered = case_when(
    registered == 1 ~ "Yes", 
    TRUE ~ "No")) %>% 
    mutate(.pred_class = case_when(
    .pred_class == 1 ~ "Yes", 
    TRUE ~ "No")) %>% 
  mutate(n_alt = paste0("(", n, ")")) %>% 
ggplot(aes(x = registered, 
           y = .pred_class, fill = n)) +
  geom_tile(color = "white") +
  geom_text(aes(label = prop), vjust = 1.5) +
  geom_text(aes(label = n_alt), vjust = 3.5) +
  scale_fill_gradient(low = "white", high = "steelblue", name = "Count") +
  labs(y = "Predicted", x = "Observed") +
  theme_minimal(base_size = 15) + 
  theme(legend.position = "none")
```

## Distribution of predictions 


```{r}
distn_fig <- predictions %>% 
  ggplot(aes(x = .pred_1)) + 
  geom_histogram(binwidth = .01, color = "black", fill = "lightblue") +
  theme_cowplot() + 
  labs(x = "Prediction", y = "count")
```



```{r}
# Train final model on the full dataset
rf_final_fit <- rf_workflow %>%
  fit(data = death_file)

# Extract feature importance
final_model <- extract_fit_parsnip(rf_final_fit)

final_importance <- vip::vi(final_model) %>% 
  mutate(Variable = case_when(
    Variable == "state" ~ "State",
    Variable == "region" ~ "Region",
    Variable == "native_language" ~ "Native Language",
    Variable == "wealth_quintile" ~ "Wealth Quintile",
    Variable == "altitude_quintile" ~ "Altitude Quintile",
    Variable == "death_age" ~ "Age at Death",
    Variable == "highest_education" ~ "Highest Education",
    Variable == "road_length_km_quantile" ~ "Road Length (km)",
    Variable == "rural" ~ "Rural Area",
    Variable == "facility_distance_quantile" ~ "Facility Distance",
    Variable == "internet" ~ "Internet Access",
    Variable == "female" ~ "Female",
    Variable == "female_head" ~ "Female Household Head",
    Variable == "health_insurance" ~ "Health Insurance",
    Variable == "cycle" ~ "Bicycle Ownership",
    Variable == "own_land" ~ "Owns Land",
    Variable == "external_death" ~ "External Cause of Death",
    TRUE ~ Variable  # Keep others unchanged
  ))


variable_importance_fig <- final_importance %>% 
  ggplot(aes(y = reorder(Variable, Importance), x = Importance)) +
  geom_point() +
  theme_cowplot() +
  geom_segment(aes(yend = Variable, x = 0, xend = Importance), linewidth = 0.5) +
  geom_segment(aes(yend = Variable, x = 0, xend = max(Importance)), linetype = "dashed", linewidth = 0.1) +
  labs(x = "MSE Difference",
       y = "")

variable_importance_fig
```


## Create combined plot 

```{r}
prediction_combined_figure <- plot_grid(confusion_matrix_plot, predicted_observed, distn_fig, variable_importance_fig)

## predictions 
ggsave(prediction_combined_figure, filename = here("figures", "prediction_combined_figure.png"), height = 10, width= 12, bg = "white") 
```

