---
title: "Machine learning set up"
author: Casey Breen
---

Other predictors: 

— Presence of educated family member
— Household size 
— Geospatial stuff
— Length of period between death and interview (months)
— Caste 
— Native language 
— Number of children under 5
— Marital status 
— Mode of transport 
— Exposure to media and literacy 
— Household member disability 
— District level covariates - district wealth, district mortality, education, etc. etc. 
— Variables for health care access 
— Digital variables - internet access, mobile phone access, etc. 


## notes for aashish 

- Missing in household and missing in 


```{r}
## library packages 
library(tidyverse)
library(haven)
library(cowplot)
library(here)
library(srvyr)
library(sl3)
library(tictoc)
library(yardstick)
library(tidymodels)

## read in death file 
death_file <- read_dta(here("data", "analysis_files", "nfhs5_deaths.DTA")) %>% 
  filter(months_since_death > 3) %>% 
  mutate(registered = as.factor(registered))
```



```{r}
death_file <- death_file %>% 
  mutate(facility_distance_quantile = ntile(facility_distance, n = 5)) %>% 
  mutate(road_length_km_quantile = ntile(road_length_km, n = 5))
  
death_file <- death_file %>% 
  filter(!is.na(area_km2))
```


```{r}
## u5 deaths 
death_file_u5 <- death_file %>% 
  filter(death_age < 5)

## 5+ deaths 
death_file_5_plus <- death_file %>% 
  filter(death_age >= 5)
```





```{r}
# set up 10 fold CV
folds_10fold_u5 <- vfold_cv(death_file_u5, 10)
folds_10fold_5_plus <- vfold_cv(death_file_5_plus, 10)


# Define the model specification
rf_model <- rand_forest(trees = 500, min_n = 5, mode = "classification") %>%
  set_engine("ranger", importance = "permutation")

# Define the recipe for data preprocessing
rf_recipe <- recipe(registered ~ death_age + female +
                      wealth_quintile + facility_distance_quantile + region + 
                      road_length_km_quantile + 
                      native_language + internet + 
                      highest_education + health_insurance + 
                      rural + own_land + cycle + 
                      external_death + female_head + state + 
                      altitude_quintile, data = death_file) %>%
  step_dummy(all_nominal_predictors()) %>% # Convert categorical to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize numeric predictors

# Create a workflow
rf_workflow <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(rf_recipe)

```

```{r}
# Fit the model
rf_fit_u5 <- rf_workflow %>%
  fit_resamples(resamples = folds_10fold_u5,
                control = control_resamples(save_pred = TRUE))

## get error metrics  
error_metrics_loco_cv_u5 <- collect_metrics(rf_fit_u5) 

## get error metrics  
predictions_u5 <- collect_predictions(rf_fit_u5) 
```


```{r}
# Fit the model
rf_fit_5plus <- rf_workflow %>%
  fit_resamples(resamples = folds_10fold_5_plus,
                control = control_resamples(save_pred = TRUE))

## get error metrics  
error_metrics_loco_cv_5_plus <- collect_metrics(rf_fit_5plus) 

## get error metrics  
predictions_5_plus <- collect_predictions(rf_fit_5plus) 
```



```{r}
# Create confusion matrix
confusion_matrix_u5 <- predictions_u5 %>%
  count(registered, .pred_class) %>%
  mutate(prop = paste0(round(n / sum(n) * 100), "%"))

# Plot with corrected code
confusion_matrix_plot_u5 <- confusion_matrix_u5 %>% 
  mutate(registered = case_when(
    registered == 1 ~ "Yes", 
    TRUE ~ "No")) %>% 
    mutate(.pred_class = case_when(
    .pred_class == 1 ~ "Yes", 
    TRUE ~ "No")) %>% 
  mutate(n_alt = paste0("(", n, ")")) %>% 
ggplot(aes(x = registered, 
           y = .pred_class, fill = n)) +
  geom_tile(color = "white") +
  geom_text(aes(label = prop), vjust = 1.5) +
  geom_text(aes(label = n_alt), vjust = 3.5) +
  scale_fill_gradient(low = "white", high = "steelblue", name = "Count") +
  labs(y = "Predicted", x = "Observed", title = "Deaths Under 5") +
  theme_minimal(base_size = 15) + 
  theme(legend.position = "none")

# Create confusion matrix
confusion_matrix_5plus <- predictions_5_plus %>%
  count(registered, .pred_class) %>%
  mutate(prop = paste0(round(n / sum(n) * 100), "%"))


# Plot with corrected code
confusion_matrix_plot_5_plus <- confusion_matrix_5plus %>% 
  mutate(registered = case_when(
    registered == 1 ~ "Yes", 
    TRUE ~ "No")) %>% 
    mutate(.pred_class = case_when(
    .pred_class == 1 ~ "Yes", 
    TRUE ~ "No")) %>% 
  mutate(n_alt = paste0("(", n, ")")) %>% 
ggplot(aes(x = registered, 
           y = .pred_class, fill = n)) +
  geom_tile(color = "white") +
  geom_text(aes(label = prop), vjust = 1.5) +
  geom_text(aes(label = n_alt), vjust = 3.5) +
  scale_fill_gradient(low = "white", high = "steelblue", name = "Count") +
  labs(y = "Predicted", x = "Observed", title = "Death 5+") +
  theme_minimal(base_size = 15) + 
  theme(legend.position = "none")
```



```{r}
age_plot <- cowplot::plot_grid(confusion_matrix_plot_u5, confusion_matrix_plot_5_plus)

ggsave(age_plot, filename = here("figures", "age_plot.png"), height = 4, width= 8, bg = "white") 
```


