---
title: "Machine learning set up"
author: Casey Breen
---

Other predictors: 

— Presence of educated family member
— Household size 
— Geospatial stuff
— Length of period between death and interview (months)
— Caste 
— Native language 
— Number of children under 5
— Marital status 
— Mode of transport 
— Exposure to media and literacy 
— Household member disability 
— District level covariates - district wealth, district mortality, education, etc. etc. 
— Variables for health care access 
— Digital variables - internet access, mobile phone access, etc. 


## notes for aashish 

- Missing in household and missing in 


```{r}
## library packages 
library(tidyverse)
library(haven)
library(cowplot)
library(here)
library(srvyr)
library(sl3)
library(tictoc)
library(yardstick)
library(tidymodels)
library(vip)         # for variable importance plots
library(geofacet)


## read in death file 
death_file <- read_dta(here("data", "analysis_files", "nfhs5_deaths.DTA")) %>% 
  filter(months_since_death > 3) %>% 
  mutate(registered = as.factor(registered)) %>% 
  group_by(state) %>%  
  filter(n() > 1000) %>% 
  ungroup()
```


```{r}
death_file <- death_file %>% 
  mutate(facility_distance_quantile = ntile(facility_distance, n = 5)) %>% 
  mutate(road_length_km_quantile = ntile(road_length_km, n = 5))
  
death_file <- death_file %>% 
  filter(!is.na(area_km2))
```



```{r}
# Define the model specification
rf_model <- rand_forest(trees = 500, min_n = 5, mode = "classification") %>%
  set_engine("ranger", importance = "permutation")

# Function to fit model for each region and extract variable importance
fit_model_by_region <- function(state_data, state_name) {
  
  # Define the recipe for data preprocessing
  rf_recipe <- recipe(registered ~ death_age + female +
                      wealth_quintile + facility_distance_quantile  + 
                      road_length_km_quantile + 
                      native_language + internet + 
                      highest_education + health_insurance + 
                      rural + own_land + cycle + 
                      external_death + female_head  + 
                      altitude_quintile, data = state_data) %>%
    step_dummy(all_nominal_predictors()) %>% 
    step_normalize(all_numeric_predictors()) 

  # Create a workflow
  rf_workflow <- workflow() %>%
    add_model(rf_model) %>%
    add_recipe(rf_recipe)

  # Fit the model
  rf_fit <- rf_workflow %>%
    fit(data = state_data %>% mutate(registered = as.factor(registered)))

  # Extract variable importance
  var_importance <- vi(rf_fit) %>%
    mutate(state = state_name) # Add state identifier

  return(list(model = rf_fit, importance = var_importance))
}

# Split training data by region
region_training_samples <- death_file %>% group_split(state)
state_names <- death_file %>% distinct(state) %>% pull(state)

# Fit models and extract variable importance
state_fits <- map2(region_training_samples, state_names, fit_model_by_region)

# Extract variable importance from each model
variable_importance_all <- bind_rows(map(state_fits, ~ .x$importance))

# View variable importance
library(forcats)

state_importance_data <- variable_importance_all %>% 
  mutate(Variable = case_when(
    Variable == "state" ~ "State",
    Variable == "region" ~ "Region",
    Variable == "native_language" ~ "Native Language",
    Variable == "wealth_quintile" ~ "Wealth Quintile",
    Variable == "altitude_quintile" ~ "Altitude Quintile",
    Variable == "death_age" ~ "Age at Death",
    Variable == "highest_education" ~ "Highest Education",
    Variable == "road_length_km_quantile" ~ "Road Length (km)",
    Variable == "rural" ~ "Rural Area",
    Variable == "facility_distance_quantile" ~ "Facility Distance",
    Variable == "internet" ~ "Internet Access",
    Variable == "female" ~ "Female",
    Variable == "female_head" ~ "Female Household Head",
    Variable == "health_insurance" ~ "Health Insurance",
    Variable == "cycle" ~ "Bicycle Ownership",
    Variable == "own_land" ~ "Owns Land",
    Variable == "external_death" ~ "External Cause of Death",
    TRUE ~ Variable  # Keep others unchanged
  )) %>% 
  mutate(state = as_factor(state)) %>% 
  mutate(state = as.character(state)) %>% 
  group_by(state) %>% 
  mutate(Importance = Importance / max(Importance))

state_importance_data %>% 
  ggplot(aes(y = state, x = Importance)) +
  ggsci::scale_fill_lancet() + 
  geom_col() + 
  theme_cowplot() + 
  theme(legend.position = "bottom") + 
  labs(x = "Importance",
       y = "") + 
  facet_wrap(~Variable)

ggsave(region_importance_plot, filename = here("figures", "region_importance_plot.png"), height = 14, width= 14, bg = "white") 
```



```{r}
state_plot <- state_importance_data %>% 
  mutate(state = str_to_title(state)) %>% 
  ggplot(aes(y = reorder(Variable, Importance), x = Importance)) + 
  geom_col() + 
  theme_cowplot() + 
  theme(legend.position = "bottom") + 
  labs(x = "Importance", y = "") + 
  facet_geo(~state, grid = india_grid1)  # Use facet_geo() properly


ggsave(state_plot, filename = here("figures", "state_plot.png"), height = 18, width= 16, bg = "white") 

```



## New version 


```{r}
# Define the model specification
rf_model <- rand_forest(trees = 500, min_n = 5, mode = "classification") %>%
  set_engine("ranger", importance = "permutation")

# Function to fit model for each region and extract variable importance
fit_model_by_region <- function(state_data, state_name) {
  
  # Define the recipe for data preprocessing
  rf_recipe <- recipe(registered ~ death_age + female +
                      wealth_quintile + facility_distance_quantile  + 
                      road_length_km_quantile + 
                      native_language + internet + 
                      highest_education + health_insurance + 
                      rural + own_land + cycle + 
                      external_death + female_head  + 
                      altitude_quintile, data = state_data) %>%
    step_dummy(all_nominal_predictors()) %>% 
    step_normalize(all_numeric_predictors()) 

  # Create a workflow
  rf_workflow <- workflow() %>%
    add_model(rf_model) %>%
    add_recipe(rf_recipe)

  # Fit the model
  rf_fit <- rf_workflow %>%
    fit(data = state_data %>% mutate(registered = as.factor(registered)))

  # Extract variable importance
  var_importance <- vi(rf_fit) %>%
    mutate(state = state_name) # Add state identifier

  return(list(model = rf_fit, importance = var_importance))
}

imp

# Rename the names of the first element of the list
# Rename the row names of the first element of the list
rownames(imp[[1]]) <- case_when(
  rownames(imp[[1]]) == "state" ~ "State",
  rownames(imp[[1]]) == "wealth_quintile" ~ "Wealth Quintile",
  rownames(imp[[1]]) == "death_age" ~ "Age at Death",
  rownames(imp[[1]]) == "native_language" ~ "Native Language",
  rownames(imp[[1]]) == "female" ~ "Female",
  rownames(imp[[1]]) == "road_length_km_quantile" ~ "Road Length (km)",
  rownames(imp[[1]]) == "highest_education" ~ "Highest Education",
  rownames(imp[[1]]) == "external_death" ~ "External Cause of Death",
  rownames(imp[[1]]) == "internet" ~ "Internet Access",
  rownames(imp[[1]]) == "cycle" ~ "Bicycle Ownership",
  rownames(imp[[1]]) == "health_insurance" ~ "Health Insurance",
  rownames(imp[[1]]) == "rural" ~ "Rural Area",
  rownames(imp[[1]]) == "female_head" ~ "Female Household Head",
  rownames(imp[[1]]) == "facility_distance_quantile" ~ "Facility Distance",
  rownames(imp[[1]]) == "own_land" ~ "Owns Land",
  TRUE ~ rownames(imp[[1]])  # Keep others unchanged
)

var_importance_plot <- plot(imp) +
  xlab("Variable importance \n (Increase in mse, test sample)") + 
  theme_cowplot()

write_rds(var_importance_plot, here("figures", "var_importance.rds"))
```

