---
title: "Machine learning set up"
author: Casey Breen
---

Other predictors: 

— Presence of educated family member
— Household size 
— Geospatial stuff
— Length of period between death and interview (months)
— Caste 
— Native language 
— Number of children under 5
— Marital status 
— Mode of transport 
— Exposure to media and literacy 
— Household member disability 
— District level covariates - district wealth, district mortality, education, etc. etc. 
— Variables for health care access 
— Digital variables - internet access, mobile phone access, etc. 


## notes for aashish 

- Missing in household and missing in 


```{r}
## library packages 
library(tidyverse)
library(haven)
library(cowplot)
library(here)
library(srvyr)
library(sl3)
library(tictoc)
library(yardstick)
library(tidymodels)

## read in death file 
death_file <- read_dta(here("data", "analysis_files", "nfhs5_deaths.DTA")) %>% 
  filter(months_since_death > 3) %>% 
  # group_by(state) %>% 
  # filter(n() > 1000) %>% 
  filter(death_age > 5) %>% 
  mutate(state = as.factor(state)) %>% 
  mutate(registered = as.factor(registered))
```



```{r}
death_file <- death_file %>% 
  mutate(facility_distance_quantile = ntile(facility_distance, n = 5)) %>% 
  mutate(road_length_km_quantile = ntile(road_length_km, n = 5))
  
death_file <- death_file %>% 
  filter(!is.na(area_km2))
```




```{r}
set.seed(12)  # Setting seed for reproducibility

# Create a random split indicator
split <- sample(1:nrow(death_file), size = 0.7 * nrow(death_file))

# Split the data into training and test sets
training_sample <- death_file[split,]
test_sample <- death_file[-split, ]
```




```{r}
## baseline model 
model <- glm(registered ~ death_age + female + as.factor(wealth_quintile) + native_language + as.factor(internet) + highest_education + as.factor(health_insurance) + as.factor(rural) + as.factor(state) + own_land + cycle + external_death + female_head + as.factor(altitude_quintile), data = training_sample, family = "binomial")

predict(model, newdata = test_sample, type = "response")

test_sample <- test_sample %>% 
  mutate(prediction_logit = predict(model, newdata = test_sample, type = "response")) 

test_sample %>% 
  mutate(prediction_binary = case_when(
    prediction_logit > 0.5 ~ 1,
    TRUE ~ 0
  )) %>% 
  count(registered, prediction_binary) %>% 
  mutate(prop = n / sum(n))
```




```{r}
# make new folds based on country 
# folds_loco <- group_vfold_cv(internet_df_train, group = state)
folds_10fold <- vfold_cv(death_file, 10)

# Define the model specification
rf_model <- rand_forest(trees = 500, min_n = 5, mode = "classification") %>%
  set_engine("ranger")

# Define the recipe for data preprocessing
rf_recipe <- recipe(registered ~ death_age + female + wealth_quintile + 
                      native_language + internet + 
                      highest_education + health_insurance + 
                      rural + own_land + cycle + 
                      external_death + female_head + 
                      altitude_quintile, data = death_file %>% mutate(state = as.character(state))) %>%
  step_dummy(all_nominal_predictors()) %>% # Convert categorical to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize numeric predictors

# Create a workflow
rf_workflow <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(rf_recipe)

# Fit the model
rf_fit <- rf_workflow %>%
  fit_resamples(resamples = folds_10fold,
                control = control_resamples(save_pred = TRUE))

## get error metrics  
error_metrics_loco_cv <- collect_metrics(rf_fit) 

## get error metrics  
predictions <- collect_predictions(rf_fit) 

death_file <- death_file %>% 
  mutate(row = row_number())

predictions <- predictions %>% 
  left_join(death_file, by = c(".row" = "row"))

predictions %>% 
  group_by(state) %>% 
  mutate(registered = as.numeric(registered.x) - 1) %>% 
  mutate(.pred_class = as.numeric(.pred_class) - 1) %>% 
  summarize(observed = mean(registered),
            predicted = mean(.pred_class),
            sample = n()) %>% 
  #filter(sample > 10) %>% 
  ggplot(aes(x = observed, y = predicted)) + 
  geom_point() + 
  ylim(0, 1) + 
  xlim(0, 1) + 
  theme_cowplot() +
  geom_abline(color = "red", linetype = "dashed") + 
  ggpubr::stat_cor()


predictions %>% 
  mutate(registered = as.numeric(registered.x) - 1) %>% 
  mutate(.pred_class = as.numeric(.pred_class) - 1) %>% 
  summarize(observed = mean(registered),
            predicted = mean(.pred_class),
            sample = n())

predictions %>% 
    mutate(registered = as.numeric(registered.x) - 1) %>% 
  group_by(female) %>% 
  mutate(decile = ntile(.pred_1, 100)) %>% 
  group_by(decile, female) %>% 
  summarize(reg = mean(registered),
            pred = mean(.pred_1)) %>% 
  ggplot(aes(x = reg, y = pred)) + 
  geom_point() + 
  geom_abline() + 
  ylim(0, 1) + 
  xlim(0, 1) + 
  theme_cowplot() + 
  facet_wrap(~female)
```

```{r}
predictions %>% 
  filter(.pred_1 > 0.95) %>% 
  count(internet)
```


```{r}
## make new folds based on country 
# folds_loco <- group_vfold_cv(internet_df_train, group = state)
folds_10fold <- group_vfold_cv(death_file, state)

# Define the model specification
rf_model <- rand_forest(trees = 500, min_n = 5, mode = "classification") %>%
  set_engine("ranger")


# Define the recipe for data preprocessing
rf_recipe <- recipe(registered ~ death_age + female + wealth_quintile + 
                      native_language + internet + 
                      highest_education + health_insurance + 
                      rural + own_land + cycle + 
                      external_death + female_head + 
                      altitude_quintile, data = death_file %>% mutate(state = as.character(state))) %>%
  step_dummy(all_nominal_predictors()) %>% # Convert categorical to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize numeric predictors

# Create a workflow
rf_workflow <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(rf_recipe)

# Fit the model
rf_fit <- rf_workflow %>%
  fit_resamples(data = death_file %>% mutate(registered = as.factor(registered)), 
                resamples = folds_10fold,
                control = control_resamples(save_pred = TRUE))

## get error metrics  
error_metrics_loco_cv <- collect_metrics(rf_res_loco_cv) 

## get error metrics  
error_metrics_loco_cv <- collect_predictions(rf_res_loco_cv) 

error_metrics_loco_cv %>% 
  ungroup() %>% 
  mutate(registered = as.numeric(registered) - 1) %>% 
  mutate(.pred_class = as.numeric(.pred_class) - 1) %>% 
  group_by(id) %>% 
  filter(n() > 300) %>% 
  summarize(observed = mean(registered),
            predicted = mean(.pred_class)) %>% 
  ggplot() + 
  geom_point(aes(x = observed, y = predicted)) + 
  ylim(0, 1) + 
  xlim(0, 1) + 
  theme_cowplot() +
  geom_abline(color = "red", linetype = "dashed")

error_metrics_loco_cv %>% 
  ungroup() %>% 
  mutate(registered = as.numeric(registered) - 1) %>% 
  mutate(.pred_class = as.numeric(.pred_class) - 1) %>% 
  # group_by(id) %>% 
  # filter(n() > 300) %>% 
  summarize(observed = mean(registered),
            predicted = mean(.pred_class)) 
```


