---
title: "Machine learning set up"
author: Casey Breen
---

Other predictors: 

— Presence of educated family member
— Household size 
— Geospatial stuff
— Length of period between death and interview (months)
— Caste 
— Native language 
— Number of children under 5
— Marital status 
— Mode of transport 
— Exposure to media and literacy 
— Household member disability 
— District level covariates - district wealth, district mortality, education, etc. etc. 
— Variables for health care access 
— Digital variables - internet access, mobile phone access, etc. 


## notes for aashish 

- Missing in household and missing in 


```{r}
## library packages 
library(tidyverse)
library(haven)
library(cowplot)
library(here)
library(srvyr)
library(sl3)
library(tictoc)
library(yardstick)
library(tidymodels)
library(vip)         

## read in death file 
death_file <- read_dta(here("data", "analysis_files", "nfhs5_deaths.DTA")) %>% 
  filter(months_since_death > 3) %>% 
  # group_by(state) %>% 
  # filter(n() > 1000) %>% 
 # filter(death_age <= 5) %>% 
  mutate(registered = as.factor(registered))
```



```{r}
death_file <- death_file %>% 
  mutate(facility_distance_quantile = ntile(facility_distance, n = 5)) %>% 
  mutate(road_length_km_quantile = ntile(road_length_km, n = 5))
  
death_file <- death_file %>% 
  filter(!is.na(area_km2))
```


```{r}
set.seed(12)  # Setting seed for reproducibility

death_file_split <- initial_split(death_file, 0.75)
training_sample <- training(death_file_split)
test_sample  <- testing(death_file_split)
```



```{r}
# Define the model specification
rf_model <- rand_forest(trees = 500, min_n = 5, mode = "classification") %>%
  set_engine("ranger", importance = "permutation")

# Function to fit model for each region and extract variable importance
fit_model_by_region <- function(region_data, region_name) {
  
  # Define the recipe for data preprocessing
  rf_recipe <- recipe(registered ~ death_age + female + wealth_quintile + 
                        native_language + internet + facility_distance_quantile + road_length_km_quantile + 
                        highest_education + health_insurance + 
                        rural  + state + own_land + cycle + 
                        external_death + female_head + 
                        altitude_quintile, data = region_data) %>%
    step_dummy(all_nominal_predictors()) %>% 
    step_normalize(all_numeric_predictors()) 

  # Create a workflow
  rf_workflow <- workflow() %>%
    add_model(rf_model) %>%
    add_recipe(rf_recipe)

  # Fit the model
  rf_fit <- rf_workflow %>%
    fit(data = region_data %>% mutate(registered = as.factor(registered)))

  # Extract variable importance
  var_importance <- vi(rf_fit$fit$fit) %>%
    mutate(region = region_name) # Add region identifier

  return(list(model = rf_fit, importance = var_importance))
}

# Split training data by region
region_training_samples <- training_sample %>% group_split(region)
region_names <- training_sample %>% distinct(region) %>% pull(region)

# Fit models and extract variable importance
region_fits <- map2(region_training_samples, region_names, fit_model_by_region)

# Extract variable importance from each model
variable_importance_all <- bind_rows(map(region_fits, ~ .x$importance))

# View variable importance
variable_importance_all %>% 
  mutate(region = as_factor(region)) %>% 
  group_by(region) %>% 
  mutate(Importance = Importance / mean(Importance)) %>% 
  ggplot(aes(y = reorder(Variable, Importance), x = Importance)) +
  geom_col(aes(fill = region), color = "black", position = position_dodge2(1)) + 
  theme_cowplot() + 
  theme(legend.position = "bottom")
```


```{r}
# Define the model specification
rf_model <- rand_forest(trees = 500, min_n = 5, mode = "classification") %>%
  set_engine("ranger", importance = "permutation")

# Define the recipe for data preprocessing
rf_recipe <- recipe(registered ~ death_age + female + wealth_quintile + facility_distance_quantile + road_length_km_quantile + 
                      native_language + region + internet + region + 
                      highest_education+ health_insurance + 
                      rural  + state + own_land + cycle + 
                      external_death + female_head + 
                      altitude_quintile, data = training_sample) %>%
  step_dummy(all_nominal_predictors()) %>% # Convert categorical to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize numeric predictors

# Create a workflow
rf_workflow <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(rf_recipe)

# Fit the model
rf_fit <- rf_workflow %>%
  fit(data = training_sample %>% mutate(registered = as.factor(registered)))

# Make predictions on the test set
test_sample <- test_sample %>%
  mutate(prediction_logit = predict(rf_fit, new_data = test_sample, type = "prob")$.pred_1)

# Convert probabilities into binary predictions
test_sample <- test_sample %>%
  mutate(prediction_binary = if_else(prediction_logit > 0.5, 1, 0))

# Summarize performance
test_sample %>%
  count(registered, prediction_binary) %>%
  mutate(prop = n / sum(n))

rf_fit %>% 
  extract_fit_parsnip() %>% 
  vip(num_features = 50)
```




