---
title: "Machine learning set up"
author: Casey Breen
---

Other predictors: 

— Presence of educated family member
— Household size 
— Geospatial stuff
— Length of period between death and interview (months)
— Caste 
— Native language 
— Number of children under 5
— Marital status 
— Mode of transport 
— Exposure to media and literacy 
— Household member disability 
— District level covariates - district wealth, district mortality, education, etc. etc. 
— Variables for health care access 
— Digital variables - internet access, mobile phone access, etc. 


## notes for aashish 

- missing in household and missing in 


```{r}
## library packages 
library(tidyverse)
library(haven)
library(cowplot)
library(here)
library(srvyr)
library(sl3)
library(tictoc)
library(yardstick)


## read in death file 
death_file <- read_dta(here("data", "analysis_files", "nfhs5_deaths.DTA")) %>% 
  filter(months_since_death > 3) %>% 
  group_by(state) %>% 
  filter(n() > 1000) %>% 
 # filter(death_age < 5) %>% 
  ungroup()
```




```{r}
death_file <- death_file %>% 
  mutate(facility_distance_quantile = ntile(facility_distance, n = 5)) %>% 
  mutate(road_length_km_quantile = ntile(road_length_km, n = 5))
  
death_file <- death_file %>% 
  filter(!is.na(area_km2))

death_file
```




```{r}
set.seed(123)  # Setting seed for reproducibility

# Create a random split indicator
split <- sample(1:nrow(death_file), size = 0.7 * nrow(death_file))

# Split the data into training and test sets
training_sample <- death_file[split, ]
test_sample <- death_file[-split, ]
```




```{r}
## baseline model 
model <- glm(registered ~ death_age + female + as.factor(wealth_quintile) + as.factor(native_language) + as.factor(internet) + highest_education + as.factor(health_insurance) + as.factor(rural) + as.factor(state) + own_land + cycle + external_death + female_head + as.factor(altitude_quintile), data = training_sample %>% filter(state == 20), family = "binomial")


test_sample <- test_sample %>% 
  mutate(prediction_logit = predict(model, newdata = test_sample %>% filter(state == 20), type = "response")) 

test_sample %>% 
  mutate(prediction_binary = case_when(
    prediction_logit > 0.5 ~ 1,
    TRUE ~ 0
  )) %>% 
  count(registered, prediction_binary) %>% 
  mutate(prop = n / sum(n))
```



```{r}
test_sample %>% 
  mutate(prediction_binary = case_when(
    prediction_logit > 0.5 ~ 1,
    TRUE ~ 0
  )) %>% 
  filter(state == 20) %>% 
  summarize(mean(prediction_binary == registered),
            mean(registered))
```



```{r}
# List of unique states
regions <- unique(training_sample$region)

# Empty list to store results for each state
state_results <- list()

# Loop over each state, fit the model, and collect the results
for (st in regions) {
  # Filter data for the specific state
  train_state <- training_sample %>% filter(region == st)
  test_state <- test_sample %>% filter(region == st)
  
  # Fit the model for the specific state
  model_state <- glm(registered ~ death_age + female + as.factor(wealth_quintile) + 
                     as.factor(internet) + highest_education + as.factor(health_insurance) + 
                     as.factor(rural) + own_land + cycle + external_death + as.factor(state) +  
                     female_head + as.factor(altitude_quintile), 
                     data = train_state, family = "binomial")
  
  # Make predictions for the test set of the specific state
  test_state <- test_state %>%
    mutate(prediction_logit = predict(model_state, newdata = test_state, type = "response")) %>%
    mutate(prediction_binary = case_when(
      prediction_logit > 0.5 ~ 1,
      TRUE ~ 0
    ))
  
  # Store the result for this state
  state_results[[st]] <- test_state %>%
    count(registered, prediction_binary) %>%
    mutate(prop = n / sum(n))
}

# Combine all state results into one dataframe for aggregation
final_results <- bind_rows(state_results, .id = "state")

# Aggregate the results across states
final_aggregated_results <- final_results %>%
  group_by(registered, prediction_binary) %>%
  summarise(n_total = sum(n), .groups = 'drop') %>%
  mutate(prop_total = n_total / sum(n_total))

# Display aggregated results
print(final_aggregated_results)
```


```{r}
# List of unique states
states <- unique(training_sample$state)

# Empty list to store results for each state
state_results <- list()

# Loop over each state, fit the model, and collect the results
for (st in states) {
  # Filter data for the specific state
  train_state <- training_sample %>% filter(state == st)
  test_state <- test_sample %>% filter(state == st)
  
  # Fit the model for the specific state
  model_state <- glm(registered ~ death_age:female + as.factor(wealth_quintile):female + 
                     as.factor(internet) + highest_education + as.factor(health_insurance) + 
                     as.factor(rural) + own_land + cycle + external_death + as.factor(district) + 
                     female_head , 
                     data = train_state, family = "binomial")
  
  # Make predictions for the test set of the specific state
  test_state <- test_state %>%
    mutate(prediction_logit = predict(model_state, newdata = test_state, type = "response")) %>%
    mutate(prediction_binary = case_when(
      prediction_logit > 0.5 ~ 1,
      TRUE ~ 0
    ))
  
  # Store the result for this state
  state_results[[st]] <- test_state 
}

# Combine all state results into one dataframe for aggregation
final_results <- bind_rows(state_results, .id = "state")

# Aggregate the results across states
final_aggregated_results <- final_results %>%
  group_by(registered, prediction_binary) %>%
  summarise(n_total = sum(n), .groups = 'drop') %>%
  mutate(prop_total = n_total / sum(n_total))

# Display aggregated results
print(final_aggregated_results)
```

```{r}
final_results %>% 
  group_by(state) %>% 
  summarize(accuracy = mean(prediction_binary == registered)) %>% 
  arrange(desc(accuracy))
```



```{r}
options(sl3.verbose = TRUE)

# Define the outcome and covariates
outcome <- "registered"
covariates <- c("death_age", "female", "wealth_quintile", "native_language", "internet",
                "highest_education", "health_insurance", "rural", "state", "own_land", 
                "cycle", "external_death", "female_head", "altitude_quintile", "facility_distance_quantile", "road_length_km_quantile")


task <- make_sl3_Task(data = training_sample, covariates = covariates, outcome = outcome, outcome_type = "binomial")

```





```{r}
# Define base learners
lrn_glm <- Lrnr_glm$new()
lrn_rf <- Lrnr_ranger$new()  # Random forest learner
lrn_gbm <- Lrnr_gbm$new()    # Gradient boosting machine learner

# Combine learners into a learner stack
learner_stack <- sl3::Stack$new(lrn_glm, lrn_rf, lrn_gbm)

```



```{r}
# Define the metalearner
metalearner <- Lrnr_nnls$new()  # Non-negative least squares (NNLS) metalearner

# Create the Super Learnerx
super_learner <- Lrnr_sl$new(learners = learner_stack, metalearner = metalearner, verbose = T)
```

```{r}
tic()
fit <- super_learner$train(task)
toc()
```


```{r}
# Create a task for the test sample
test_task <- make_sl3_Task(data = test_sample, covariates = covariates, outcome = outcome, outcome_type = "binomial",)

# Make predictions
predictions <- fit$predict(test_task)

# Add predictions to test sample
test_sample <- test_sample %>%
  mutate(prediction = predictions) %>%
  mutate(prediction_binary = ifelse(prediction > 0.5, 1, 0))

# Evaluate performance
result <- test_sample %>%
  count(registered, prediction_binary) %>%
  mutate(prop = n / sum(n))

print(result)

```



```{r}
test_sample %>% 
  group_by(state) %>% 
  count(registered == prediction_binary)
```

